### 2014-12-08 tex/short_2015/body.tex
more text

\title[Automatic Proxy Geometry]%
      {Automatic Proxy Geometry}

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
\author[J.\,O. Nygaard \& J. Mikkelsen]
       {J.\,O. Nygaard$^{1}$
        and J. Mikkelsen$^{1}$
        and Husk Anonymisering Hvis Paakrevet!$^{1}$
%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
        \\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
         $^1$SINTEF ICT, applied mathematics, Norway
%        $^2$ Another Department to illustrate the use in papers from authors
%             with different affiliations
       }

@@

% \teaser{
%  \includegraphics[width=\linewidth]{eg_new}
%  \centering
%   \caption{New EG Logo}
% \label{fig:teaser}
% }

@@

\begin{abstract}
   We describe an algorithm and implementation of an automatic proxy geometry
   system for a server/client remote rendering setup. The client may be a thin
   client in the form of a hand-held pad, for our implementation we only assume
   the availability of WebGL and Javascript. For our rendering server, we make
   use of a GL-capable web server. In situations where the rate of received
   images deteriorate due to low bandwith, high latency or long rendering times,
   the client displays an automatically generated proxy geometry. This proxy
   geometry is computed from depth buffers bundled with rendered images from the
   server, and requires no modification of the application or its data itself.
   % Husk: Skal vaere en blank linje nedenfor

\begin{classification} % according to http://www.acm.org/class/1998/
\CCScat{FIX THIS Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

@@

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered (OpenGL)]{
    \includegraphics[width=.47\linewidth]{frview1.jpg}
  }
  \subfigure[Client-rendered proxy model (Google Chrome)]{
    \includegraphics[width=.47\linewidth]{frview2.jpg}
  }
  \caption{\label{fig:FRView} The oil reservoir viewer FRView, showing a
  server-rendered image and a proxy geometry generated from this and the
  accompanying depth buffer. Note that we have chosen parameters for the proxy
  generation (small number of splats ($128^2$), forcing display of splats that
  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
  the ragged edges, a few holes, and the oddly placed (and rendered) axis
  indicator.}
\end{figure}

@@

One effect of this, also enhanced by new cloud-based approaches, wireless
connection, and mobile devices with display capabilities outpacing their
computation and storage capabilities, is that interactivity still is a difficult
issue. In this work, we consider a server/client model for rendering, but this
is just a way of separating the interaction from the rendering; the discussion
also applies to a more integrated setup.  In order to obtain the frame rates on
the client suitable for interactive visualizations, demands are put on
bandwidth, latency and rendering speed. If one or more of these demands are not
met, according to a suitable measure, the frame rate will deteriorate together
with the user experience.

One way to deal with such situations is for the client to have a {\em
proxy model}, defined as a temporary model to be shown and manipulated
locally on the client while waiting for the appropriate image from the server.
A particularly unwelcomed chore is that of building proxy geometries for various
data sets. The typical ``hello world'' proxy geometry is the familiar bounding
box wire frame, but this will degrade the user experience unacceptably for most
applications. One alternative is to hand craft such simpler proxies. They have
to be simple both for the building cost, and also for the actual rendering on
the possibly very thin client. Another alternative is to automate this process,
a large triangulation may for instance be represented by a much coarser
approximating triangulation that can be built automatically. One disadvantage of
this approach is that this is not easy to do for many kinds of data, like our
main case in which we have an oil reservoir viewer (FRView) that renders large
{\em corner points grids}, together with {\em faults}, oil wells, and more. See
Figure~\ref{fig:FRView} above for an example of both a fully completed
rendering, and an automatically generated proxy geometry rendered in Google
Chrome.

@@

We also mention that even if bandwidth and latency is not a problem, it may be
desirable to let a server of limited capacity serve many simultaneous users,
hence decreasing the effective server time available for each one. In this case,
suitable scaling properties may still be achieved without increasing server
capacity. \textbf{referanse til ArrowHead her?}

\subsection{Previous work}

Our approach has some similarities to {\em image based rendering} (IBR)
techniques (\textbf{referanse til noe av dette?}), with the difference that an
important IBR problem would be the reconstruction of a depth map from images,
while we have access to the full depth map from a rendering pass.
Another way to use depth maps similar to what we do, is for ``immersive
streaming'',
see \eg, \texttt{http://www.cc...},
% http://www.cc.gatech.edu/conferences/3DPVT08/Program/Papers/paper203.pdf
where focus is on depth map compression, an issue that we also must confront.

@@

@@

The main idea is to automatically construct a proxy geometry on the client side
from depth data provided by the server. For a new image to be displayed on the
client, the server renders an image in a frame buffer, and as part of this
process it also renders a corresponding depth buffer. (This means that our
technique is not immediately ready for \eg, ray-tracing.) This depth buffer may
be regarded as a height map over a viewing plane relative to the observing
client. Hence, this depth map does not contain information about scene elements
not in view for the set of view parameters at the time of server side
rendering. Our approach is based on the assumption that small transformations of
this height map still will give good approximations to the scene being
rendered. These renderings are exactly what the client will compute and show
while waiting for the next update from the server. See
Figure~\ref{fig:2Dheightmap} and Figure~\ref{fig:2DheightmapRotated} below for an
illustration of this heightmap computed on the server, and a recovered 3D model
on the client.

\begin{figure}[htb]
  \centering
  \input{fig2.tex}

  \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
           height map seen from above, for a certain transformation. Also shown
           are the corresponding RGB-fragments (red and green) and depth
           fragments (floating point scalars shown as grayscale pixels.)}
\end{figure}

@@

Further compression methods (\textbf{passende med en ref her?}) may be employed
to bring this data size down even more, but then the cost of the compression
itself must also be added to the cost of rendering on the server.
%
%The cost must of course be considered if the operation of packaging images for
%sending to the client is significant, which may be the case if, \eg, images must
%be compressed or converted to some special format. An example of the latter
%might be {\texttt base64}-encoding for converting the image to a purely ASCII
%format.
%
Together with the rendered rgb image and the depth image we will also send the
current view transformations back to the client.

@@

As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
height map does not really allow us more than concluding where a set of points
belong to the 3D model, these points giving rise to exactly the depth map from
the server. Given these points, the client may build a model approximating the
actual scene being rendered on the server. Since the information from such a
depth map can only contain the foremost point along any ray from the observer
toward the scene, it is often said to be in 2.5D, as opposed to purely 2D or
fully 3D.

@@

It makes sense for the client only to render at most one 3D point for each depth
fragment available. This may be too much for a thin client, and one may want to
operate with a smaller number of 3D points, and instead render each of them with
a larger number of pixels on the client side, this is called a {\em
splat}. A set of such splats for a given depth image, we will call a {\em proxy
model}. There are some issues to discuss in relation to this, \eg, what size and
shape of geometries to render for each 3D point, how to color them, how to
compute depth fragments and so on.

@@

A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. For the splat centered in
$\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be centered on the client's canvas at
screen coordinate $\pv_{i, j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i,
j}$, the texture coordinate transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{\text{splats}_x} & 0 \\
    0 & \frac{1}{\text{splats}_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{\text{vpwidth}}{\text{splats}_x} & 0 \\
     0 & \frac{\text{vpheight}}{\text{splats}_y}
  \end{pmatrix}
  \cdot \text{overlap},
\]
\[
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B
  \cdot \text{overlap},
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ together with the
``overlap'' just provides a scaling factor so that ``overlap=1'' produces
non-overlapping but covering splats when $M_c=M_s$, \ie, the client does no
transformation of the proxy model.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We compute $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions and
performing the perspective division and transformation into window coordinates
in the vertex shader. Let proxy model positions corresponding to three corners
of a quadrilateral splat in clip coordinates be
given by
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer and
transformed to $[-1, 1]$. By combining this with a $w$-component set to one, we
have in effect done a perspective division, so that the right-hand side vertices
are in clip space, suitable for multiplication with $U$.

We use
\begin{equation}
  \Delta x = \text{splats}_x / \text{DepthImg-width}
  \label{eq:deltaChoice}
\end{equation}
\[
  \Delta y = \text{splats}_y / \text{DepthImg-height}.
\]
\textbf{fiks de likningene over}
Note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
chosen something larger. The values are used for looking up depth image samples,
so we should not choose something too small, our choice is exactly to look for
neighbouring depth samples. Since this calculation of $\sv_x$ and $\sv_y$ is
really a difference approximating a gradient on the geometry, we should not make
$\Delta x$ and $\Delta y$ to large either. Experiments show us
that~(\ref{eq:deltaChoice}) is really a good choice.

This then leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and after perspective division and viewport transformation, we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, which we can subtract to obtain 2D vectors spanning the splat,
\[
  \sv_{\Delta x} =
  \sv_{x+\Delta x} - \sv =
    \frac{\text{width}}{2\Delta x} \left(
        \frac{\pv_{x+\Delta x}.xy}{\pv_{x+\Delta x}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
and
\[
  \sv_{\Delta y} =
  \sv_{y+\Delta y} - \sv =
    \frac{\text{height}}{2\Delta y} \left(
        \frac{\pv_{y+\Delta y}.xy}{\pv_{y+\Delta y}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
where we have used the ``shader notation'' to extract 2D vectors and scalars
from the 4D vectors, and $\text{width}$ and $\text{height}$ are the viewport
width and height, respectively.

Then, on the client we render a \texttt{glPoint} for each splat, with texture
coordinates $(s, t)^\prime$ and \texttt{glPointSize} set to a larger value that will
produce many fragments. Each fragment will then look up the server-rendered
image at position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
\]
where $(u, v)^\prime = \texttt{gl\_PointCoord}-0.5$ are ``intra-splat texture
coordinates''. \textbf{Boer kanskje sjekke at det ikke er noen faktor 2 som har
falt ut noe sted paa veien...}

@@

@@

As soon as the client has a set of such proxy models, it will also need an
algorithm for replacing old models with new ones.  One can imagine a plethora of
{\em splat set replacement algorithms}, we have tested three approaches that all
retain a constant number of proxy models. The first algorithm is to replace the
one with a viewing direction differing the most from the newly received
model. The second simply replaces the oldest one in store. The third replaces a
proxy model $j$ if the replacement results in the following objective function
being reduced,
\[
  \text{coverage}_j = 
  \sum_{i=1, i\neq j}^n 
    \angle(\textbf{camdir}_i, \textbf{new camdir} )^2,
\]
where $\textbf{camdir}$ is the direction in which the camera was looking for the
generation of that particular proxy model.

There are some issues to consider with respect to panning and zooming that we
will not discuss here due to the limited space available.  All of these model
replacement algorithms work decently well.

%-------------------------------------------------------------------------
\subsection{Dynamic/heuristic method selection, depth compression}

Since the main purpose of the automatically generated proxy model is to
facilitate a client-rendered image while waiting for one from the server for
increased interactivity, it is important that the process of generating and
sending proxy model data from the server itself is not hold up more than
necessary. There are mainly three sources of delay (\textbf{passer aa nevne
tidligere?}) for server-rendered images to the client; high
latency, low bandwidth, and slow server-rendering itself.

@@

To take care of this, we have adopted an adaptive specification of
proxy model data from the server. This involves a more light-weight image (lossy
JPG-compression with adaptive quality control) while the user is interacting
with the client, as well as a reduced resolution of the depth buffer sent from
the server. See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
high-quality proxy models.

\begin{figure}[htb]
  \centering
  \subfigure[Full png image,\newline full depth buffer]{
    \includegraphics[width=.30\linewidth]{auto1.jpg}
  }
  \subfigure[Coarse jpg,\newline quality $q=0$]{
    \includegraphics[width=.30\linewidth]{auto2.jpg}
  }
  \subfigure[Auto-proxy, trun-\newline cated depth data]{
    \includegraphics[width=.30\linewidth]{auto3.jpg}
  }
  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
           available bandwidth and server capacity. The server-rendered JPG
           image adaptively selects a quality setting based on recorded
           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
           the last step before switching to ``auto proxy'' (c). The auto-proxy
           model shown displays artifacts caused both by the magenta surface
           being almost parallel to the viewing direction at the moment of proxy
           model recording, and by the non-smooth parts of the geometry
           (corners). The number of splats is set low on purpose, and the depth
           buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$.}
\end{figure}

@@

The automatic proxy geometry implementation is fully invisible to the
application, meaning that all existing applications immediately will have the
feature available. We have tested the algorithm on several smaller test cases,
but also on a larger oil reservoir viewer, FRView, developed in the CloudViz
project, see~\cite{cloudviz}.

This viewer utilizes several GLSL shaders to render reservoir cells and
boundaries ({\em corner point grids}), tubular wells etc., and operates in
general on a geometric model that is not trivial to reduce in complexity. The
model is typically also very large, meaning that rendering the full model on a
thin client is prohibitive. With the automatic proxy geometry, we can obtain
interactive frame rates with a limited connection from a very lightweight client
to the server. For a comparison of a server-rendered image and a client-rendered
proxy model that is slightly rotated on the client, see Figure~\ref{fig:FRView}.

@@

There are several directions in which we would like to follow up and improve
this concept,
\begin{itemize}
\item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
all relevant viewing parameters on the client, with the help of a minimal set of
proxy models.
\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
envision a spatially partioned grid, for instance containing a distance field,
with the geometry as an iso-surface.
\end{itemize}

%-------------------------------------------------------------------------
\section{Midlertidige notater}

\begin{itemize}

>>>>

%\title[Automatic Proxy Geometry Generation]%
%      {Automatic Proxy Geometry Generation}

\title[Automatic Proxy Model Generation]%
      {Hybrid Client/Server Rendering with Automatic Proxy Model Generation}

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
\author[J.\,O. Nygaard \& J. Mikkelsen]
       {J.\,O. Nygaard$^{1}$
        and J. Mikkelsen$^{1}$
        and Husk Anonymisering Hvis Paakrevet! \textbf{JA DET ER DET}$^{1}$
%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
        \\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
%         $^1$SINTEF ICT, applied mathematics, Norway
%        $^2$ Another Department to illustrate the use in papers from authors
%             with different affiliations
       }

@@

\teaser{
  \centering
  \subfigure[Server-rendered (OpenGL)]{
    \includegraphics[width=.2\linewidth]{frview1.jpg}
  }
  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts not subdued)]{
    \includegraphics[width=.2\linewidth]{frview2.jpg}
  }
  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts subdued)]{
    \includegraphics[width=.2\linewidth]{frview2.jpg}
  }
  \caption{\label{fig:FRView} The oil reservoir viewer FRView, developed in the
  CloudViz project~\cite{cloudviz}, showing a server-rendered image and two
  client-rendered images of automatically generated proxy models. For~(b),
  parameters are chosen to highlight some artifacts (few and large splats
  without intra-splat texturing, displaying of skewed splats) while the second
  client-rendered image (c) has parameters chosen to make an optimal proxy
  geometry (many more splats, intra-splat texturing, discarding of skewed
  splats.)}
}

@@

\begin{abstract}
   We describe an algorithm and implementation of an automatic proxy model
   generation system for a client/server remote rendering setup. In this
   context, a {\em proxy model}, is a light-weight replacement of a main model
   that is much easier to manipulate and render. This proxy model is
   automatically derived from the main model. The client may be a {\em thin
   client}, \eg, a hand-held tablet. For our implementation we only assume the
   availability of WebGL and Javascript on this client. On the server side we
   make use of OpenGL and a web server. In situations where the rate of received
   images deteriorate due to low bandwith, high latency or long rendering times,
   the client displays the proxy model. The proxy geometry is computed from
   depth buffers bundled with rendered images from the server, and the algorithm
   is non-intrusive in that it requires no modification of the application or
   its data itself.
   % Husk: Skal vaere en blank linje nedenfor

\begin{classification} % according to http://www.acm.org/class/1998/
\CCScat{\textbf{FIX THIS}Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

@@

@@

One effect of this, also enhanced by new cloud-based approaches, wireless
connection, and mobile devices with display capabilities outpacing their
computation and storage capabilities, is that interactivity still is a difficult
issue. In this work, we consider a client/server model for rendering, but this
is just a way of separating the interaction from the rendering; the discussion
also applies to a more integrated setup.  In order to obtain the frame rates on
the client suitable for interactive visualizations, demands are put on
bandwidth, latency and rendering speed. If one or more of these demands are not
met, according to a suitable measure, the frame rate will deteriorate together
with the user experience.

%\begin{figure}[htb]
%  \centering
%  \subfigure[Server-rendered (OpenGL)]{
%    \includegraphics[width=.47\linewidth]{frview1.jpg}
%  }
%  \subfigure[Client-rendered proxy model (Google Chrome)]{
%    \includegraphics[width=.47\linewidth]{frview2.jpg}
%  }
%  \caption{\label{fig:FRView} The oil reservoir viewer FRView \textbf{FIKS HER ref}, showing a
%  server-rendered image and a proxy geometry generated from this and the
%  accompanying depth buffer. Note that we have chosen parameters for the proxy
%  generation (small number of splats ($128^2$), forcing display of splats that
%  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
%  the ragged edges, a few holes, and the oddly placed (and rendered) axis
%  indicator. \textbf{Flytte opp til teaser, tredje bilde: uten artifacter for aa
%  vise at det kan bli bra}}
%\end{figure}

One way to deal with such situations is for the client to have a {\em proxy
model}, defined as a temporary model to be shown and manipulated locally on the
client while waiting for the appropriate image from the server.  A particularly
unwelcomed chore is that of building proxy geometries for various data sets. The
typical ``hello world'' proxy geometry is the familiar bounding box wire frame,
but this will degrade the user experience unacceptably for most
applications. One alternative is to hand craft such simpler proxies. They have
to be simple both for the building cost, and also for the actual rendering on
the possibly very thin client. Another alternative is to automate this process,
a large triangulation may for instance be represented by a much coarser
approximating triangulation that can be built automatically. One disadvantage of
this approach is that this is not easy to do for many kinds of data, like our
main case in which we have an oil reservoir viewer (FRView,~\cite{cloudviz})
that renders large {\em corner point grids}, together with faults, oil wells,
and more. See Figure~\ref{fig:FRView} above for an example of both a fully
completed rendering, and automatically generated proxy geometries rendered in
Google Chrome.

@@

We also mention that even if bandwidth and latency is not a problem, it may be
desirable to let a server of limited capacity serve many simultaneous users,
hence decreasing the effective server time available for each one. In this case,
suitable scaling properties may still be achieved without increasing server
capacity. The results described in this paper will be commercialized in the new
SINTEF project \mbox{ArrowHead}~\cite{arrowhead}.

\subsection{Previous work}
\label{sec:prevWork}

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider.

Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could also in principle do this for our 2.5D maps fetched from
the server, but it is not clear that the benefit would outweight the
cost. \marginpar{\textbf{REF?!}}

@@

Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization. % , see~\eg,~\cite{}
\marginpar{\textbf{REF?!}}

@@

The main idea is to automatically construct a proxy geometry on the client side
from depth data provided by the server. The server already produces a depth
buffer as part of its rendering process.  This means that our technique is not
immediately ready for \eg, ray-tracing.) This depth buffer may be regarded as a
height map over a viewing plane relative to the observing client. Hence, this
depth map does not contain information about scene elements not in view for the
set of view parameters at the time of server side rendering. Our approach is
based on the assumption that small transformations of this height map still will
give good approximations to the scene being rendered. These renderings are
exactly what the client will compute and show while waiting for the next update
from the server. See Figure~\ref{fig:2Dheightmap} and
Figure~\ref{fig:2DheightmapRotated} below for an illustration of this heightmap
computed on the server, and a recovered 3D model on the client.

\begin{figure}[htb]
  \centering
  \input{fig2.tex}
  \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
           height map seen from above, for a certain transformation. Also shown
           are the corresponding RGB-fragments (red and green) and depth
           fragments (floating point scalars shown as grayscale pixels.)}
\end{figure}

@@

Further compression may bring this data size down even more, but the cost of
compression/decompression must also be added to the cost of rendering on the
client/server. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
proposed that promises to be fast both for the compression and decompression
stages. \marginpar{\textbf{REF?!}} Together with the rendered rgb image and the
depth image we will also send the current view transformations back to the
client.

@@

\marginpar{\textbf{boer kunne krympes}}
As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
height map does not really allow us more than concluding where a set of points
belong to the 3D model, these points giving rise to exactly the depth map from
the server. Given these points, the client may build a model approximating the
actual scene being rendered on the server. Since the information from such a
depth map can only contain the foremost point along any ray from the observer
toward the scene, it is often said to be in 2.5D, as opposed to purely 2D or
fully 3D.

@@

It makes sense for the client only to render at most one 3D point for each depth
fragment available. This may be too much for a thin client, and one may want to
operate with a smaller number of 3D points, and instead render each of them with
a larger number of pixels on the client side, this is called a {\em splat}. A
set of such splats for a given depth image is what we have termed a {\em proxy
model}. There are some issues to discuss in relation to this, \eg, what size and
shape of geometries to render for each 3D point, how to color them, how to
compute depth fragments and so on.  By rendering splats, we get fewer ``false
connections'' than if we connect 3D points and reconstruct topology, for
instance in the shape of a triangulation. On the other hand, we risk getting
more ``holes'' in our models.

@@

A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
projection matrices on the client and server, respectively, and $M_c$ and $M_s$
corresponding view matrices. Let there be $n_x \times n_y$ splats laid out in a
grid. For the splat centered in $\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be
centered on the client's canvas at screen coordinate $\pv_{i, j} = P_c M_c
M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{n_x} & 0 \\
    0 & \frac{1}{n_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{\text{vpwidth}}{n_x} & 0 \\
     0 & \frac{\text{vpheight}}{n_y}
  \end{pmatrix}
  \cdot \text{overlap},
\]
\[
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B
  \cdot \text{overlap},
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ together with the
``overlap'' just provides a scaling factor so that ``overlap=1'' produces
non-overlapping but covering splats when $M_c=M_s$, \ie, the client does no
transformation of the proxy model.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We compute $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions and
performing the perspective division and transformation into window coordinates
in the vertex shader. Let proxy model positions corresponding to three corners
of a quadrilateral splat in clip coordinates be
given by
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer
$\text{depthTex}$ and transformed to $[-1, 1]$. By combining this with a
$w$-component set to one, we have in effect done a perspective division, so that
the right-hand side vertices are in clip space, suitable for multiplication with
$U$.

We use
\begin{equation}
  \begin{aligned}
    \Delta x & = n_x / \text{width}(\text{depthTex}) \\
    \Delta y & = n_y / \text{height}(\text{depthTex}),
  \end{aligned}
  \label{eq:deltaChoice}
\end{equation}
note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
chosen something larger. The values are used for looking up depth image samples,
so we should not choose something too small, our choice is exactly to look for
neighbouring depth samples. Since this calculation of $\sv_x$ and $\sv_y$ is
really a difference approximating a gradient on the geometry, we should not make
$\Delta x$ and $\Delta y$ to large either. Experiments show us
that~(\ref{eq:deltaChoice}) is really a good choice.

This then leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and after perspective division and viewport transformation, we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, which we can subtract to obtain 2D vectors spanning the splat,
\[
  \sv_{\delta} =
  \sv_{x+\delta} - \sv =
    \frac{L}{2\delta} \left(
        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
where $\delta$ is either $\Delta x$ or $\Delta y$, $L$ is either
$\text{width}$ or $\text{height}$, and we have used the ``shader
notation'' to extract 2D vectors and scalars from the 4D vectors, and
$\text{width}$ and $\text{height}$ are the viewport width and height,
respectively.

Then, on the client we render a \texttt{glPoint} for each splat, with texture
coordinates $(s, t)^\prime$ and \texttt{glPointSize} set to a larger value that will
produce many fragments. Each fragment will then look up the server-rendered
image at position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
  =
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    \texttt{glPointCoord}-\frac{1}{2}
  \end{pmatrix}
\]
where $(u, v)$ are ``intra-splat texture coordinates''.

@@

%-------------------------------------------------------------------------
\subsubsection{Other splat considerations}

\marginpar{\textbf{... samle avsnittene nedenfor og komprimere...}}

@@

As soon as the client has a set of such proxy models, it will also need an
algorithm for replacing old models with new ones.  One can imagine a plethora of
{\em splat set replacement algorithms}, we have tested three approaches that all
retain a constant number of proxy models. The first algorithm is to replace the
one with a viewing direction differing the most from the newly received
model. The second simply replaces the oldest one in store. The third replaces a
proxy model $k$ if the replacement results in the following objective function
being reduced,
\[
  \text{coverage} =
  \sum_{i=1}^n 
    \sum_{j\neq i}^n 
      ( \angle(\textbf{camdir}_i, \textbf{camdir}_j )^2,
\]
where $n$ is the number of proxy models available and the model to be replaced
is the one that minimizes this after replacement.
%\textbf{droppe fra her: }
%\ie, if $\text{coverage}_k < \text{coverage}$, where
%\[
%  \text{coverage}_k =
%  \min_j
%  \sum_{i=1, i\neq j}^n 
%    ( \angle(\textbf{camdir}_i, \textbf{new\ camdir}_j )^2.
%\]
%\textbf{til hit?}
Here, $\textbf{camdir}$ is the direction in which the camera was looking for the
generation of that particular proxy model.  There are some issues to consider
with respect to panning and zooming that we will not discuss here due to the
limited space available.

%-------------------------------------------------------------------------
\subsection{Dynamic/heuristic method selection, depth compression, auto-tuning}
\label{sec:autoTuning}

Since the main purpose of the automatically generated proxy model is to
facilitate a client-rendered image while waiting for one from the server for
increased interactivity, it is important that the process of generating and
sending proxy model data from the server itself is not hold up more than
necessary. There are mainly three sources of delay for server-rendered images to
the client; high latency, low bandwidth, and slow server-rendering itself.

@@

To take care of this, we have adopted an adaptive specification of
proxy model data from the server. This involves a more light-weight image (lossy
JPG-compression with adaptive quality control) while the user is interacting
with the client, as well as a reduced resolution of the depth buffer sent from
the server. Another possible form of auto-tuning is to let the client adaptively
choose the number of splats, stored number of proxy models, etc.

%See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
%high-quality proxy models.
%\begin{figure}[htb]
%  \centering
%  \subfigure[Full png image,\newline full depth buffer]{
%    \includegraphics[width=.30\linewidth]{auto1.jpg}
%  }
%  \subfigure[Coarse jpg,\newline quality $q=0$]{
%    \includegraphics[width=.30\linewidth]{auto2.jpg}
%  }
%  \subfigure[Auto-proxy, trun-\newline cated depth data]{
%    \includegraphics[width=.30\linewidth]{auto3.jpg}
%  }
%  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
%           available bandwidth and server capacity. The server-rendered JPG
%           image adaptively selects a quality setting based on recorded
%           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
%           the last step before switching to ``auto proxy'' (c). The auto-proxy
%           model shown displays artifacts caused both by the magenta surface
%           being almost parallel to the viewing direction at the moment of proxy
%           model recording, and by the non-smooth parts of the geometry
%           (corners). The number of splats is set low on purpose, and the depth
%           buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$. \textbf{KRYMPE HER}}
%\end{figure}

@@

The automatic proxy geometry implementation in Tinia is fully invisible to the
application, meaning that all existing applications immediately will have the
feature available. Without Tinia, the algorithm is minimally intrusive in that
only the depth buffer will have to be added to the rendering output of the
application. We have tested the algorithm on several smaller test cases, but
also on a larger oil reservoir viewer, FRView, developed in the CloudViz
project, see~\cite{cloudviz}.

This viewer utilizes several GLSL shaders to render reservoir cells and
boundaries ({\em corner point grids}), tubular wells etc., and operates in
general on a geometric model that is not trivial to reduce in complexity. The
model is typically also very large, meaning that rendering the full model on a
thin client is prohibitive. With the automatic proxy geometry, we can obtain
interactive frame rates with a limited connection from a very lightweight client
to the server. For a comparison of a server-rendered image and a client-rendered
proxy model that is slightly rotated on the client, see again Figure~\ref{fig:FRView}.

@@

There are several directions in which we would like to follow up and improve
this concept, foremost of these are,
\begin{itemize}

\item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
all relevant viewing parameters on the client, with the help of a minimal set of
proxy models.

%\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
%envision a spatially partioned grid, for instance containing a distance field,
%with the geometry as an iso-surface.
